New theoretical lower bounds of weak Schur numbers using regular Schur numbers

### 2021-06-04 article/article.tex
Correction numérotation
corrections
accolades
elsarticle
adieu sat
ptites modifs

\documentclass{article}
\title{New lower bounds for Schur and weak Schur numbers}
\author{Romain Ageron, Paul Casteras, Thibaut Pellerin, Yann Portella}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}

@@

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{nicematrix}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

\newtheorem{definition}{Definition}[section]
\newtheorem{notation}[definition]{Notation}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{computational theorem}{Computational Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}[section]
\newtheorem{proposition}{Proposition}[section]

@@

@@

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. Finding templates allows us to 
get explicit partitions improving lower bounds as well as the growth rate for both Schur numbers and Ramsey 
numbers \(R_n(3)\). We also developed a new method to improve lower bounds for weak Schur numbers. 
Furthermore, this paper tries to analyze former works on weak Schur numbers based on the principle that good 
partitions into \(n+1\) weakly sum-free subsets start with a good partition into \(n\) weakly sum-free subsets. 
We show that exceeding the previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption 
upon imposing certain conditions on the good 5-subsets partitions. The new lower bounds include 
\(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS (9) \geqslant 22\,536 \) and 
\(\WS (10) \geqslant 71\,214 \).

\include{introduction.tex}

% \include{definitions_notations.tex}

\include{schur.tex}

\include{weakschur.tex}

\include{search_space.tex}

>>>>

\documentclass[3p]{elsarticle}
\title{New lower bounds for Schur and weak Schur numbers}
\author{Romain Ageron, Paul Casteras, Thibaut Pellerin, Yann Portella}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{MyMnSymbol}

@@

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{nicematrix}

@@

\newtheorem{definition}{Definition}[section]
\newtheorem{notation}[definition]{Notation}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{computational theorem}[definition]{Computational Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{remark}[definition]{Remark}
\newtheorem{proposition}[definition]{Proposition}

@@

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. Finding templates allows us to 
get explicit partitions improving lower bounds as well as the growth rate for both Schur numbers and Ramsey 
numbers \(R_n(3)\). We also developed a method to improve lower bounds for weak Schur numbers. 
Furthermore, we try to analyze former works on weak Schur numbers based on the principle that \textit{good} 
partitions into \(n+1\) weakly sum-free subsets start with a \textit{good} partition into \(n\) weakly sum-free subsets. 
We show that exceeding the previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption 
upon imposing certain conditions on the \textit{good} 5-subsets partitions. The new lower bounds include 
\(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS (9) \geqslant 22\,536\) and 
\(\WS (10) \geqslant 71\,214 \).

@@

\input{introduction.tex}

\input{schur.tex}

\input{weakschur.tex}

### 2021-06-04 rapport/introduction.tex
ptites modifs

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
This article focuses on the lower bounds for these numbers.

@@

\qquad Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It used to give the best lower bounds for all sufficiently large numbers. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not utilize the weakly sum-free hypothesis.

@@

\par
In 2020, Rowley introduced the notion of template for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and gives new lower bounds (and inequalties) for Schur numbers. 
Rowley also gives two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds which besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\vspace{1.5ex}
* denotes an exact value, not just a lower bound
\end{center}

@@

\qquad The main contribution of this article is a generalization of the concept of template to weak Schur numbers. 
It gives new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a special 
case an analogous for weak Schur numbers of Abbott and Hanson's construction for Schur numbers.

\par
In the \hyperref[Schur]{first section}, we explain Rowley's template-based construction in the context of 
Schur numbers and then give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
Then, in the \hyperref[WeakSchur]{second section}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.

\par
Finally, in the \hyperref[SearchSpace]{third section}, we analyze the significant difference between new lower 
bounds obtained with templates and the former lower bounds obtained by computer search and we provide 
evidence which indicate that the main assumption made in those articles removes the optimal partitions from the 
search space.

>>>>

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
Our study focuses on the lower bounds for these numbers.

@@

\qquad Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It used to give the best lower bounds for all sufficiently large numbers. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not use the weakly sum-free hypothesis.

@@

\par
In 2020, Rowley introduced the notion of templates for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and produces new lower bounds (and inequalties) for Schur numbers. 
He also provides two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds which besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\vspace{1.5ex}
* denotes the exact value
\end{center}

@@

\qquad Our main contribution is a generalization of the concept of template to weak Schur numbers. 
Our templates provide new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a special 
case an analogous for weak Schur numbers of Abbott and Hanson's construction for Schur numbers.

\par
In the \hyperref[Schur]{section 2}, we explain Rowley's template-based construction in the context of 
Schur numbers and then give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
Then, in the \hyperref[WeakSchur]{section 3}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.

\par
Finally, in the \hyperref[SearchSpace]{section 4}, we analyze the significant difference between new lower 
bounds obtained with templates and the former lower bounds obtained by computer search and we provide 
evidence which indicates that the main assumption made in those articles removes the optimal partitions from the 
search space.

### 2021-06-04 rapport/schur.tex
ptites modifs

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there exists a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) is the minimum of the special subset in the \(SF\)-template.
\end{theorem}

@@

	\par
	NB: In the following 3 predicates, the conditions \(x + y \leqslant q\)  and \(x + y \leqslant p\) are omitted for readability.
	\par
	The sum-free condition is expressed as
	\[\forall (x,y) \in [\![1,q]\!]^2, f(x) = f(y) \Longrightarrow f(x+y) \neq f(x)\]
	\[\forall (x,y) \in [\![1,p]\!]^2, g(x) = g(y) \Longrightarrow g(x+y) \neq g(x)\]

	The additionnal constraint for the SF-template is
	\[
	\forall (x,y) \in [\![1,q]\!]^2, \left\{
	\begin{array}{l}
		f(x) = f(y) \leqslant n \\
		x + y > q
	\end{array}
	\right. \Longrightarrow f(x+y-q) \neq f(x)
	\]

	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), write \(x = (\alpha - 1) + u\) for some integers \(\alpha \in \mathbb{Z}\) and 
	\(u \in [\![1,q]\!]\). This decomposition is of course unique. \(\alpha\) (resp. \(u\)) can be interpreted as the row 
	(resp. column) number of \(x\). Define a new coloring \(h\) as follows
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!] \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u) & \text{if}~f(u) \leqslant n \\
			n + g(\alpha) & \text{if}~f(u) = n + 1 \\
		\end{array} \right.
	\end{array}
	\]

	\(h\) is well-defined since, by definition of \(m_{n+1}\), \(\forall x \in [\![p q + 1, p q + m_{n + 1} - 1 ]\!], f(u) 
	\leqslant n\) and therefore \(\forall x \in [\![1,pq+m_{n+1}-1]\!], f(u) = n + 1 \implies \alpha \in [\![1, p]\!]\).

	\par
	We now prove that \(h\) is a sum-free coloring. Let \(x,y \in [\![1,pq + m_{n+1}-1]\!]\) such that \(h(x) = h(y)\) 
	and \(x+y \leqslant pq+m_{n+1}-1\). We claim that \(h(x+y) \neq h(x)\). Write \(x = (\alpha - 1) q + u\) and 
	\(y = (\beta - 1) q + v\) where \(\alpha, \beta \in \mathbb{Z}\) and \(u, v \in [\![1,q]\!]\). Distinguish between 
	two cases according to the value of \(h(x)\). \\

	\noindent \underline{\textbf{Case 1:} \(h(x) \leqslant n\)}
	\par
	Assume that \(h(x+y) \leqslant n\), otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition of \(h\) 
	and given that \(h(u) = h(v)\), \(f(u) = f(v)\). Distinguish between two cases according to the value of \(x + y\).

	\begin{itemize}
	\item If \(u + v > q\), write \(w = u + v - q \in [\![1, q]\!]\). Then \(x + y = (\alpha + \beta - 1) q + w\). By definition, 
		\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant n\), the additionnal constraint on \(f\) implies \(f(w) 
		\neq f(u)\), that is \(h(x + y) \neq h(x)\).
	\item If \(u + v \leqslant p\), write \(w = u + v \in [\![1, q]\!]\). Then \(x+y = (\alpha + \beta- 2) q + w\). By definition, 
		\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant n\), the sum-free property of \(f\) implies \(f(w) \neq f(u)\), 
		that is \(h(x + y) \neq h(x)\).
	\end{itemize}

	\noindent \underline{\textbf{Case 2:} \(h(x) \geqslant n + 1\)}
	\par
	\(h(x) = n + g(\alpha) = 
	k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
	to the value of \(x + y\).

@@

The following proposition can help improve the additive constant of a SF-template. Although it does not allow us to 
improve the SF-templates we have found, the analogous of this proposition for WSF-templates (see next section) 
allows us to improve one of them.

\begin{proposition}
Let \((q, n) \in \mathbb{N}^*)^2\) and let \(f\) be a coloring associated to a SF-template with width \(q\) and \(n+1\) 
colors. Let \(b \in \mathbb{N}\) (\(b = m_{n+1} - 1\) works) and assume there there exists a coloring \(g\) of 
\([\![1, b]\!]\) with \(n+1\) colors such that

\begin{itemize}
\item \(\forall (x, y) \in [\![1, q]\!]^2, \left\{
	\begin{array}{l}
		f(x) = f(y) \\
		(x + y) \mod q \leqslant b
	\end{array}
	\right. \implies g((x + y) \mod q) \neq f(x)\)
\item \(\forall (x, y) \in [\![1, q]\!] \times  [\![1, b]\!],  \left\{
	\begin{array}{l}
		f(x) = g(y) \\
		x + y \leqslant b
	\end{array}
	\right. \implies g(x + y) \neq f(x)\)
\end{itemize}

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - p S(n))\), we have
\[ S(n+k) \geqslant S^+(n+1)S(k) + b\]
\end{proposition}

@@

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a SF-template with width \(q\) and \(n+1\) colors,
	and a SF-template with width \(p\) and \(k\) colors, then there exists SF-template with width \(pq\) and \((n+k)\) 
	colors.
	\end{theorem}

	And the associated inequality

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S^+(n+k) \geqslant S^+(n+1)S^+(k) \]
\end{corollary}

@@

We exhibited SF-templates using a SAT solver, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We have sought templates providing the greatest value of 
\((a, b)\) (for the lexicographic order). When the number of colors excceeded 5, in order to reduce the search space we 
only looked for symmetric SF-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

@@

The first inequality comes from  Schur's original article\cite{Schur1917}. The second one is due to
Abott \cite{AbbottHanson}
and the third one to Rowley \cite{RowleyRamsey}. The other ones are new.

\par
The first 3 inequalities are optimal. The fourth one is optimal among symmetric SF-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimum. 
Finally, the sixth one is obtained by combining a SF-template with width 380 and one with width 3. 
Although we could not find a better SF-template with 7 colors, the last inequality is definitely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

>>>>

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there exists a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element in the special subset in the \(SF\)-template.
\end{theorem}

@@

	\par
	NB: In the following three predicates, the conditions \(x + y \leqslant q\)  and \(x + y \leqslant p\) are omitted for readability.
	\par
	The sum-free condition is expressed as:
	\[\forall (x,y) \in [\![1,q]\!]^2, f(x) = f(y) \Longrightarrow f(x+y) \neq f(x)\],
	\[\forall (x,y) \in [\![1,p]\!]^2, g(x) = g(y) \Longrightarrow g(x+y) \neq g(x)\].

	The additionnal constraint for the SF-template is:
	\[
	\forall (x,y) \in [\![1,q]\!]^2, \left\{
	\begin{array}{l}
		f(x) = f(y) \leqslant n \\
		x + y > q
	\end{array}
	\right. \Longrightarrow f(x+y-q) \neq f(x)
	\].

	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), write \(x = (\alpha - 1) + u\) for certain integers \(\alpha \in \mathbb{Z}\) and 
	\(u \in [\![1,q]\!]\). This decomposition is of course unique. \(\alpha\) (resp. \(u\)) can be interpreted as the row 
	(resp. column) number of \(x\). A new coloring \(h\) is defined as follows:
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!], \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u), & \text{if}~f(u) \leqslant n, \\
			n + g(\alpha), & \text{if}~f(u) = n + 1. \\
		\end{array} \right.
	\end{array}
	\]

	Function \(h\) is well-defined since, by definition of \(m_{n+1}\), \(\forall x \in [\![p q + 1, p q + m_{n + 1} - 1 ]\!], f(u) 
	\leqslant n\) and therefore \(\forall x \in [\![1,pq+m_{n+1}-1]\!], f(u) = n + 1 \implies \alpha \in [\![1, p]\!]\).

	\par
	We now prove that \(h\) is a sum-free coloring. Let \(x,y \in [\![1,pq + m_{n+1}-1]\!]\) such that \(h(x) = h(y)\) 
	and \(x+y \leqslant pq+m_{n+1}-1\). We claim that \(h(x+y) \neq h(x)\). We write \(x = (\alpha - 1) q + u\) and 
	\(y = (\beta - 1) q + v\) where \(\alpha, \beta \in \mathbb{Z}\) and \(u, v \in [\![1,q]\!]\). Two cases are to be
	distinguished according to the value of \(h(x)\). \\

	\noindent \underline{\textbf{Case 1:} \(h(x) \leqslant n\)}
	\par
	Let us assume that \(h(x+y) \leqslant n\), otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition of function 
	\(h\) and given that \(h(u) = h(v)\), we conclude \(f(u) = f(v)\). Two cases are to b distinguished according to the value of \(x + y\).

	\begin{itemize}
	\item If \(u + v > q\), we write \(w = u + v - q \in [\![1, q]\!]\). Consequentely \(x + y = (\alpha + \beta - 1) q + w\). By definition, 
		\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant n\), the additionnal constraint on \(f\) implies \(f(w) 
		\neq f(u)\), that is \(h(x + y) \neq h(x)\).
	\item If \(u + v \leqslant p\), we write \(w = u + v \in [\![1, q]\!]\). Consequentely \(x+y = (\alpha + \beta- 2) q + w\). By definition, 
		\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant n\), the sum-free property of \(f\) implies \(f(w) \neq f(u)\), 
		that is \(h(x + y) \neq h(x)\).
	\end{itemize}

	\noindent \underline{\textbf{Case 2:} \(h(x) \geqslant n + 1\)}
	\par
	Now we have \(h(x) = n + g(\alpha) = 
	k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
	to the value of \(x + y\).

@@

The following proposition may improve the additive constant of a SF-template. Although it does not allow us to 
improve the SF-templates we have found, the analogous of this proposition for WSF-templates (see next section) 
allows us to improve one of them.

\begin{proposition}
Let \((q, n) \in \mathbb{N}^*)^2\) and let \(f\) be a coloring associated to a SF-template with width \(q\) and \(n+1\) 
colors. Let \(b \in \mathbb{N}\) and assume there is a coloring \(g\) of 
\([\![1, b]\!]\) with \(n+1\) colors such that:

\begin{itemize}
\item \(\forall (x, y) \in [\![1, q]\!]^2, \left\{
	\begin{array}{l}
		f(x) = f(y) \\
		(x + y) \mod q \leqslant b
	\end{array}
	\right. \implies g((x + y) \mod q) \neq f(x)\),
\item \(\forall (x, y) \in [\![1, q]\!] \times  [\![1, b]\!],  \left\{
	\begin{array}{l}
		f(x) = g(y) \\
		x + y \leqslant b
	\end{array}
	\right. \implies g(x + y) \neq f(x)\).
\end{itemize}

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - p S(n))\), we have
\[ S(n+k) \geqslant S^+(n+1)S(k) + b\].
\end{proposition}

@@

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there is a SF-template with width \(q\) and \(n+1\) colors,
	and a SF-template with width \(p\) and \(k\) colors, then there also is a SF-template with width \(pq\) and \((n+k)\) 
	colors.
	\end{theorem}

	The inequality associated with this theorme is given by:

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S^+(n+k) \geqslant S^+(n+1)S^+(k) \].
\end{corollary}

@@

We produced SF-templates using a SAT solver, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing the largest value of 
\((a, b)\) (in the lexicographic order). When the number of colors excceeded five, in order to reduce the search space we 
only looked for symmetric SF-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

@@

The first inequality comes from  Schur's original article\cite{Schur1917}. The second one is due to
Abott \cite{AbbottHanson}
and the third one to Rowley \cite{RowleyRamsey}. The three last inequalies are our result.

\par
The first three inequalities are optimal. The fourth one is optimal among symmetric SF-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimum. 
Finally, the sixth one is obtained by combining a SF-template with width 380 and one with width 3. 
Although we could not find a better SF-template with seven colors, the last inequality is definitely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

### 2021-06-04 rapport/weakschur.tex
ptites modifs

\par
NB: In the following 5 predicates, the conditions \(x + y \leqslant p\)  and \(x + y \leqslant a + b\) are omitted for readability.
\par
The (weakly) sum-free conditions are expressed as 
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	f(x) = f(y) \\
	x \neq y
\end{array}
\right. \Longrightarrow f(x+y) \neq f(x)
\]
\[
\forall (x,y) \in [\![b+1,a + b]\!]^2, f(x) = f(y) \Longrightarrow f(x+y) \neq f(x)
\]
\[
\forall (x,y) \in [\![1,p]\!]^2, g(x) = g(y) \Longrightarrow g(x+y) \neq g(x)
\]

The additionnal constraints for the WSF-template are
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	f(x) = f(y) \leqslant n \\
	x + y > a + b
\end{array}
\right. \Longrightarrow f(\pi(x+y)) \neq f(x)
\]
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	f(x) = f(y) = n + 1 \\
	x + y > 2 a + b
\end{array}
\right. \Longrightarrow f(x+y - 2 a) \neq f(x)
\]

Here, we consider \(\pi\) defined in the previous subsection as an application defined on \([\![b + 1, p a + b]\!]\). 
Split \([\![1, p a + b]\!]\) into 3 subsets.

@@

Define a new coloring \(h\) as follows
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto & 
	\left\{ \begin{array}{l l}
		f(x) & \text{if}~x \in \mathcal{T} \\
		f(\pi(x)) & \text{if}~x \in \mathcal{C} \\
		n + g(\lambda(x)) & \text{if}~x \in \mathcal{R}
	\end{array} \right.
\end{array}
\]

\(h\) is well defined since \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) is a partition of  \([\![1, p a + b]\!]\). 
We now prove that \(h\) is a weakly sum-free coloring. Let \(x,y \in [\![1, p a + b]\!]\) such that \(x \neq y\), 
\(h(x) = h(y)\) and \(x+y \leqslant p a+ b\). We claim that \(h(x+y) \neq h(x)\). Distinguish between six cases 
according to the subsets \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) to which \(x\) and \(y\) belong. It is 
sufficient to check only six cases out of nine since \(x\) and \(y\) play symmetric roles. \\

>>>>

\par
NB: To keep the notation short, the conditions \(x + y \leqslant p\)  and \(x + y \leqslant a + b\) are omitted in the following five predicates.
\par
The (weakly) sum-free conditions are expressed as:
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	f(x) = f(y) \\
	x \neq y
\end{array}
\right. \Longrightarrow f(x+y) \neq f(x)
\]
\[
\forall (x,y) \in [\![b+1,a + b]\!]^2, f(x) = f(y) \Longrightarrow f(x+y) \neq f(x),
\]
\[
\forall (x,y) \in [\![1,p]\!]^2, g(x) = g(y) \Longrightarrow g(x+y) \neq g(x).
\]

The additionnal constraints for the WSF-template are:
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	f(x) = f(y) \leqslant n \\
	x + y > a + b
\end{array}
\right. \Longrightarrow f(\pi(x+y)) \neq f(x),
\]
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	f(x) = f(y) = n + 1 \\
	x + y > 2 a + b
\end{array}
\right. \Longrightarrow f(x+y - 2 a) \neq f(x).
\]

Here, we consider \(\pi\) defined in subsection 3.2 \textbf{LABEL NECESSAIRE} as an application defined on \([\![b + 1, p a + b]\!]\). 
Split \([\![1, p a + b]\!]\) into three subsets.

@@

A new coloring \(h\) is defined as follows:
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto & 
	\left\{ \begin{array}{l l}
		f(x) & \text{if}~x \in \mathcal{T} \\
		f(\pi(x)) & \text{if}~x \in \mathcal{C} \\
		n + g(\lambda(x)) & \text{if}~x \in \mathcal{R}
	\end{array} \right.
\end{array}
\]

Function \(h\) is well defined since \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) is a partition of  \([\![1, p a + b]\!]\). 
We now prove that \(h\) is a weakly sum-free coloring. Let \(x,y \in [\![1, p a + b]\!]\) be such that \(x \neq y\), 
\(h(x) = h(y)\) and \(x+y \leqslant p a+ b\). We claim that \(h(x+y) \neq h(x)\). We distinguish between six cases 
according to the subsets \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) to which \(x\) and \(y\) belong. It is 
sufficient to check only six cases out of nine since \(x\) and \(y\) play symmetric roles. \\
